### 2023-01-02 writeups/Observations.tex
increase test dataset size
add k-fold xvalidation
edit observations
run friedman tests on model outputs

For example, for the counter \texttt{\{"big": 50, "brown": 25, "fox": 25\}}, the antisemitic dataset would have a counter \texttt{\{"big": 0.056159, "brown": 0, "fox": 0\}}, and the baseline would have \texttt{\{"big": 0.125219, "brown": 0, "fox": 0.008348\}}. The next step would be to sort these words alphabetically, but luckily the order they come in in the text is already alphabetical. Next, they would be represented as vectors and the distance would be found between them and the input text, as shown:\\
\[
\left[ {\begin{array}{c}
    50\\
    25\\
    25\\
  \end{array} } \right] - \left[ {\begin{array}{c}
    0.056159\\
    0\\
    0\\
  \end{array} } \right] = \left[ {\begin{array}{c}
    49.943841\\
    25\\
    25\
  \end{array} } \right] = \vec{a} , \left| \vec{a} \right| \approx 61.19
\]

\[
\left[ {\begin{array}{c}
    50\\
    25\\
    25\\
  \end{array} } \right] - \left[ {\begin{array}{c}
    0.125219\\
    0\\
    0.008348\\
  \end{array} } \right] = \left[ {\begin{array}{c}
    49.874781\\
    25\\
    24.991652\\
  \end{array} } \right] = \vec{b},\hfill \left| \vec{b} \right| \approx 61.13
\]

@@

Three machine learning models were used: SVMs, Decision Trees, and Naïve Bayes classifiers. Each of those models was used for Boolean and type classification and with three seperate text representations (Bag-of-Words, raw frequency, and TF-IDF), making a total of 18 models. For the sake of evaluation, the rules-based model was retrained on only the training data, then tested on the test data. Models labeled RB are rules-based. The scores represent the proportion of the test dataset that the model correctly predicted, 1 means perfect predictions, 0 means entirely incorrect predictions.
\begin{center}
Table 3: Machine learning model evaluations and comparisons to rules-based model
\begin{tabular}{ |c|c|c|c| }
\hline {\bf Model type} & {\bf Text representation} & {\bf Boolean score} & {\bf Type score} \\
\hline Decision Tree & BoW & 0.653 &  0.513\\
\hline Decision Tree & Frequency & 0.695  &  0.692\\
\hline Decision Tree & TF-IDF & 0.703 & 0.487\\
\hline SVM & BoW & 0.703 & 0.692\\
\hline SVM & Frequency & 0.703 & 0.564\\
\hline SVM & TF-IDF & 0.678 & 0.462\\
\hline Naïve Bayes & BoW & 0.669 & 0.769\\
\hline Naïve Bayes & Frequency & 0.602 & 0.487\\
\hline Naïve Bayes & TF-IDF & 0.602 & 0.462\\
\hline Cosine Similarity (RB) & Frequency & 0.602 & 0.462\\
\hline Vector Distance (RB) & Frequency & 0.602 & 0.462\\
\hline Vector Distance (RB) & TF-IDF & 0.602 & 0.462\\
\hline
\end{tabular}
\end{center}
\end{document}

>>>>

For example, for the counter \texttt{\{"big": 50, "brown": 25, "fox": 25\}}, the antisemitic dataset would have a counter \texttt{\{"big": 0.056159, "brown": 0, "fox": 0\}}, and the baseline would have \texttt{\{"big": 0.125219, "brown": 0, "fox": 0.008348\}}. The next step would be to sort these words alphabetically, but luckily the order they come in in the text is already alphabetical. Next, they would be represented as vectors and the distance would be found between them and the input text, as shown:\\
\[
\left[ {\begin{array}{c}
    50\\
    25\\
    25\\
  \end{array} } \right] - \left[ {\begin{array}{c}
    0.056159\\
    0\\
    0\\
  \end{array} } \right] = \left[ {\begin{array}{c}
    49.943841\\
    25\\
    25\
  \end{array} } \right] = \vec{a} , \left| \left| \vec{a} \right| \right| \approx 61.19
\]

\[
\left[ {\begin{array}{c}
    50\\
    25\\
    25\\
  \end{array} } \right] - \left[ {\begin{array}{c}
    0.125219\\
    0\\
    0.008348\\
  \end{array} } \right] = \left[ {\begin{array}{c}
    49.874781\\
    25\\
    24.991652\\
  \end{array} } \right] = \vec{b},\hfill \left| \left| \vec{b} \right| \right| \approx 61.13
\]

@@

Three machine learning models were used: SVMs, Decision Trees, and Naïve Bayes classifiers. Each of those models was used for Boolean and type classification and with three seperate text representations (Bag-of-Words, raw frequency, and TF-IDF), making a total of 18 models. For the sake of evaluation, the rules-based model was retrained on only the training data, then tested on the test data. Models labeled RB are rules-based. The scores represent the proportion of the test dataset that the model correctly predicted, 1 means perfect predictions, 0 means entirely incorrect predictions. The models were evaluated using 10-fold cross-validation. Tables with rankings for the specific algorithms and text representations are given after.

\begin{center}
Table 3: Machine learning model evaluations --- binary classification
\begin{tabular}{ |c|c|c| }
\hline {\bf Model type} & {\bf Text representation} & {\bf Score}\\
\hline Decision Tree & BoW & 0.722\\
\hline Decision Tree & Frequency & 0.756\\
\hline Decision Tree & TF-IDF & 0.760\\
\hline SVM & BoW & 0.769\\
\hline SVM & Frequency & 0.757\\
\hline SVM & TF-IDF & 0.732\\
\hline Naïve Bayes & BoW & 0.638\\
\hline Naïve Bayes & Frequency & 0.591\\
\hline Naïve Bayes & TF-IDF & 0.599\\
\hline
\end{tabular}
\end{center}

\begin{center}
Table 4: Algorithm comparison --- binary classification\\
\begin{tabular}{ |c|c| }
\hline {\bf Model type}& {\bf Score}\\
\hline Support Vector Machine & 0.714\\
\hline Decision Tree & 0.568\\
\hline Naïve Bayes & 0.466\\
\hline
\end{tabular}
\end{center}

\begin{center}
Table 5: Representation comparison --- binary classification\\
\begin{tabular}{ |c|c| }
\hline {\bf Representation}& {\bf Score}\\
\hline TF-IDF & 0.550\\
\hline Frequency & 0.516\\
\hline Bag of Words & 0.507\\
\hline
\end{tabular}
\end{center}

Using a Friedman Chi Squared test, the p-value for table 3 is $6.30 \times 10^{-89}$, the p-value for the different algorithms is $4.39 \times 10^{-73}$, and the p-value for the different representations is $5.21 \times 10^{-10}$.

\begin{center}
Table 6: Machine learning model evaluations --- type classification\\
\begin{tabular}{ |c|c|c| }
\hline {\bf Model type} & {\bf Text representation} & {\bf Score} \\
\hline Decision Tree & BoW & 0.576\\
\hline Decision Tree & Frequency & 0.567\\
\hline Decision Tree & TF-IDF & 0.555\\
\hline SVM & BoW & 0.582\\
\hline SVM & Frequency & 0.521\\
\hline SVM & TF-IDF & 0.476\\
\hline Naïve Bayes & BoW & 0.515\\
\hline Naïve Bayes & Frequency & 0.538\\
\hline Naïve Bayes & TF-IDF & 0.526\\
\hline
\end{tabular}
\end{center}

\begin{center}
Table 7: Algorithm comparison --- type classification\\
\begin{tabular}{ |c|c| }
\hline {\bf Model type}& {\bf Score}\\
\hline Support Vector Machine & 0.479\\
\hline Decision Tree & 0.321\\
\hline Naïve Bayes & 0.300\\
\hline
\end{tabular}
\end{center}

\begin{center}
Table 8: Representation comparison --- type classification\\
\begin{tabular}{ |c|c| }
\hline {\bf Representation}& {\bf Score}\\
\hline TF-IDF & 0.397\\
\hline Bag of Words & 0.347\\
\hline Frequency & 0.338\\
\hline
\end{tabular}
\end{center}

\end{document}
